### 2012-11-10 checkin/project-checkin.tex
checkin

\documentclass[letterpaper,12pt]{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage{enumerate}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{float}
\usepackage[usenames]{xcolor}

\title{MATH-469 Project Checkin}
\author{Eric Biggers}

\begin{document}
\maketitle

The goal of my project is to implement an algorithm that can assemble a genome,
given a set of reads that came from it.  The key idea in the algorithm I am
implementing is the construction of a string graph where the edges are labeled
with DNA sequences and the vertices correspond to reads, and a path through the
graph corresponds to a set of reads that assemble together consistently.

I have implemented the initial stages of the algorithm.  Each step has been
implemented as a separate binary program.  This means that the overall assembly
process will iteratively transform the data (including the graph) on-disk into
the final assembly.  The currently implemented programs are:

\begin{itemize}
\item
	{\tt convert-reads}:  Imports and merges a set of reads
	into a concise binary representation.
\item
	{\tt compute-overlaps}:  Takes as input a set of reads and outputs all
	pairwise overlaps between reads of a minimum length specified by the
	{\tt --min-overlap-len} option.  The algorithm avoids a pairwise
	comparison of all reads by using a hash table to find reads sharing a
	seed.  Only exact overlaps are considered at this point.

\item
	{\tt print-overlaps}:  Prints a textual representation of a set of
	overlaps, given a file containing the overlaps in binary format.

\item
	{\tt remove-contained-reads}:
	Given a set of reads and the overlaps that were computed from them,
	finds all reads that are fully contained by another read and discard
	them, along with the corresponding overlaps.  If a pair of reads is
	identicaly, only one copy of the read is kept.

\item
	{\tt build-directed-string-graph}:
	Given a set of reads and the overlaps that were computed from them,
	build the {\bf directed} string graph that models the genome assembly.
\item
	{\tt build-bidirected-string-graph}:
	Given a set of reads and the overlaps that were computed from them,
	build the {\bf bidirected} string graph that models the genome assembly.

\item
	{\tt print-string-graph}:
	Prints a textual representation of a directed or bidirected string
	graph, or prints statistics about the graph.

\item
	{\tt transitive-reduction}:
	Given a string graph, remove edges $v \to x$ where there exist edges $v
	\to w \to x$, provided that $v \to x$ is labeled by the same sequence as
	that of $v \to w$ concatenated with $v \to x$.  This program currently
	only works on directed string graphs, but the bidirected case is in
	progress.

\item
	{\tt collapse-unbranched-paths}:
	Given a string graph, collapse chains of vertices that have in-degree 1
	and out-degree 1.  This program currently only works on directed string
	graphs, but the bidirected case is in progress.

\item
	{\tt digraph-to-bidigraph}:
	Convert a directed string graph into a bidirected string graph.

\end{itemize}

The modular design allows for multiple possible data flows, but a possible flow
is {\tt convert-reads} $\to$ {\tt compute-overlaps} $\to$ {\tt
remove-contained-reads} $\to$ \newline {\tt build-directed-string-graph} $\to$ {\tt
transitive-reduction} $\to$ {\tt digraph-to-bidigraph}, which will produce a
transitively-reduced, collapsed bidirected string graph produced from overlaps
between non-contained reads.

Even with the final stages not implemented, the algorithm is already effective
when used on random genomes because with no spurious overlaps, the
transitively-reduced string graph is a chain of vertices with no branching,
which then collapses into a single edge in {\tt collapse-unbranched-paths}.
This is true even if the genome is 15 million random base pairs, given 100 bp
reads sampled uniformly from either strand with no errors.  (Such a genome takes
about 1 minute to assemble with maximum memory usage 1.8 GB, with the main time
and memory bottleneck being the {\tt compute-overlaps} program).

What follows is an explanation of the initial stages of the algorithm on a very
simple case.  The diagram below shows a 64- base-pair ``genome'' that has had
three 32-base-pair error-free reads at a constant separation of 12 base pairs.
Note that the genome is two-stranded.  {\tt A} always pairs with {\tt T} and
{\tt C} always pairs with {\tt G}, but the two strands run in opposite
directions.  Reads 1 and 2 come from the top (reverse strand), while read 3
comes from the bottom (forward strand).  The reads are directed as shown (e.g.
Read 1 begins with {\tt ACAGC}, reading from right to left, not {\tt TGAGC}).

%make out.{bi,}digraph.{pdf,dot} out.reduced.{bi,}digraph.pdf
%out.reduced.collapsed.{bi,}digraph.pdf SAMPLE_SIZE=64 USE_PIRS=false
%MIN_OVERLAP_LEN=8 READ_LEN=32 READ_SEP=12 -B
\vspace{0.5cm}
\nopagebreak

\noindent
{\tt 1.}
{\tt \color{red} $\leftarrow$TGAGCTTAAGGCTTATCTATCTTCAGACGACA$\leftarrow$ }\\
{\tt 2.}
\hspace*{2.58cm}{\tt \color{red} $\leftarrow$TTATCTATCTTCAGACGACTATTATAGCGCGG$\leftarrow$}\\
{\tt G.}
{\tt $\leftarrow$TGAGCTTAAGGCTTATCTATCTTCAGACGACTATTATAGCGCGGCCAAGACTACGCGGAGCCCC$\leftarrow$} \\
{\tt G.}
{\tt $\rightarrow$ACTCGAATTCCGAATAGATAGAAGTCTGCTGATAATATCGCGCCGGTTCTGATGCGCCTCGGGG$\rightarrow$} \\
{\tt 3.}
\hspace*{5.16cm}{\tt \color{blue} $\rightarrow$TCTGCTGATAATATCGCGCCGGTTCTGATGCG$\rightarrow$}

From looking at the diagram, it is apparent that Read 1 overlaps with Read 2 by
20 base pairs, Read 1 overlaps with Read 2 by 20 base pairs, and Read 1 overlaps
with Read 3 by 8 base pairs.  These overlaps are computed by the {\tt
compute-overlaps} program.  Note that an overlap may match either the forward or
reverse-complement sequence.

Before building the string graph, {\tt remove-contained-reads} is run to remove reads
that are fully contained by another read.  There are no such reads in this
example.

Given the set of overlaps, the next step is to build a bidirected or directed
string graph.  Both these graphs are shown on the next page.  In the directed
graphs, a node $n.B$ means the beginning of read $n$, while the node $n.E$ means
the end of read $n$.

Consider graph (2A), the directed string graph after transitive reduction, and
consider the walk $1.B \to 2.B \to 3.E$.  This represents a way in which reads
1, 2, and 3 assemble together consistently.  In the simplest case, all the nodes
in a path would be $E$ nodes, and this would correspond to a sequence of reads
that assemble together as they are, with no reverse-complementation.  However,
in this case, we walk from the {\it beginning} of read 1 to the {\it beginning}
of read 2, indicating that they are laid out in reverse-complement order, before
reaching the {\it end} of read 3, which is in forward order.  This
reconstruction is consistent with how the reads were actually sampled from the
genome (shown earlier).

The other component of graph (2A) produces an equivalent assembly, but it lays
out the reads in the opposite order and assembles the other strand of the
genome.

Graph (2B) is equivalent to (2A), but combines every node $n.B$ and $n.E$ into
one node $n$ to deal with the double-strandedness in a more concise manner.  The
edges are bidirected.  A bidirected edge may be traversed either way in a walk,
but if a vertex is entered through an inward head, it must be left through an
outward head, and vice versa.

In both the directed and bidirected graphs, the edges are labeled by DNA
sequences, although in the images, only the lengths of the sequences are shown.

\newpage
\begin{tabular}{|p{0.5\textwidth}|p{0.5\textwidth}|} \hline
\includegraphics[scale=0.6]{out.digraph-crop.pdf}
\newline 1A. Directed string graph of original data
&
\includegraphics[scale=0.6]{out.bidigraph-crop.pdf}
\newline 1B. Equivalent bidirected string graph to (1A)
\\ \hline

\includegraphics[scale=0.6]{out.reduced.digraph-crop.pdf}
\newline 2A. Graph (1A) after transitive reduction.  Read 2 is between reads 1
and 3, so there is no need for edges directly between 1 and 3.
&
\includegraphics[scale=0.6]{out.reduced.bidigraph-crop.pdf}
\newline 2B. Graph (1B) after transitive reduction, or the equivalent bidirected
string graph to (2A)

\\ \hline

\includegraphics[scale=0.6]{out.reduced.collapsed.digraph-crop.pdf}
\newline 3A. Graph (2A) after collapsing unbranched paths
&
\includegraphics[scale=0.6]{out.reduced.collapsed.bidigraph-crop.pdf}
\newline 3B. Graph (2B) after collapsing unbranched paths, or the equivalent
bidirected string graph to (3A).

\\ \hline

\end{tabular}
\newpage

%\begin{minipage}[c]{0.4\linewidth}
%\includegraphics[scale=0.6]{out.digraph-crop.pdf}
%\captionof{figure}{Directed string graph of original data}
%\end{minipage}
%\begin{minipage}[c]{0.4\linewidth}
%\includegraphics[scale=0.6]{out.bidigraph-crop.pdf}
%\captionof{figure}{Equivalent bidirected string graph of original data}
%\end{minipage}

%\hrule

%\begin{minipage}{0.5\linewidth}
%\includegraphics[scale=0.6]{out.reduced.digraph-crop.pdf}
%\end{minipage}
%\begin{minipage}{0.5\linewidth}
%\includegraphics[scale=0.6]{out.reduced.bidigraph-crop.pdf}
%\end{minipage}

%\hrule

%\begin{minipage}{0.5\linewidth}
%\includegraphics[scale=0.6]{out.reduced.collapsed.digraph-crop.pdf}
%\end{minipage}
%\begin{minipage}{0.5\linewidth}
%\includegraphics[scale=0.6]{out.reduced.collapsed.bidigraph-crop.pdf}
%\end{minipage}

\end{document}