### 2014-03-24 docs/hw6/hw6.tex
Cleanup repository

\documentclass[letterpaper,12pt]{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage{enumerate}

\title{MATH-469 Homework 6}
\author{Eric Biggers}

\begin{document}
\maketitle

The goal of my project is to implement an algorithm that can assemble a genome,
given a set of reads that came from it.  For the purposes of this project, a
genome is defined as a sequence of A's, T's, C's, and G's, along with the
corresponding opposite strand that matches T's with A's and G's with C's.  The
input to the algorithm is a set of {\em reads} that are substrings from the
genome.  Each read may be from either the ``forward'' strand of the genome or
the ``reverse'' strand of the genome, but the reads will always be directed
forwards along the ``forward'' strand or backwards along the ``reverse'' strand
(this corresponds to the sequencing reactions always proceeding from the 5' to
the 3' end of each strand).

The reads will not be required to be of any particular length, other than long
enough to actually compute some overlaps, but the longer they are, the better
the algorithm should work.  The algorithm will not account for any pairing
information that may exist among reads (e.g. reads that came from the same DNA
fragment or insert of known size).  Also, I will initially assume that the reads
do not contain any sequencing errors and that there is no foreign DNA in the
sample.

The overall description of the algorithm is as follows:

\begin{enumerate}
\setcounter{enumi}{-1}
\item For testing the algorithm, the initial step is to simulate reads from a
sample genome.  An existing program can be used for this.
\item Compute all overlaps of some length $\ell$ among the reads.  An overlap is
where the bases at one end of a read match the bases at one end of another read.
The bases may match literally, or they may match the reverse-complement
sequence.  The length $\ell$ is chosen so that the probability of two random DNA
sequences having an overlap of length $\ell$ purely by chance is sufficiently
low; this will depend on the genome size.

The original algorithm accounts for erroneous data by looking not for perfect
overlaps, but rather for overlaps where the edit distance between the overlapped
regions is less than a certain fraction of the overlap.  (i.e., the bases might
be different by up to a certain threshold, such as 2.5\%, and still be
considered a valid overlap).  I will initially work with error-free data and
compute only perfect overlaps, but this could be something to look into
implementing later.

The naive approach to computing overlaps requires comparing all $N$ of the reads
to every other read--- an $O(N^2)$ approach that would likely be a significant
bottleneck in the overall algorithm.  I plan to speed up the running time of
this step of the algorithm by using a seed-and-extend approach where only reads
sharing a $k$-mer are compared, where $k \le \ell$ (for perfect overlaps).  This
can be done by placing each $k$-mer, along with its location, in a hash table
indexed by the $k$-mer.

\item Build a graph where each vertex corresponds to either the beginning or the
end of a read, and the overlaps are used to connect vertices.  (Edges are
labeled with DNA sequences.)  The exact edges that should be added to the graph
given an overlap $o$ are described in the original paper.  Basically, the edges
are always labeled with sequences oriented {\em outwards} from the overlapped
region, and this sequence may be either forward or reverse-complement.  Reads
that are fully contained by another read are not included in the graph, as they
theoretically do not contain any information not provided by the longer read.
The main challenge in this step is probably to represent the graph in a
memory-efficient way.
\item Delete all transitive edges from each vertex (i.e. edges $f \to h$ where
there are also edges $f \to g \to h$), since these edges are redundant.  The
algorithm for this step is provided in the original paper, and pseudocode is
provided, so it should be fairly easy to implement this step, although again it
might be tricky to keep memory usage from being too high.
\item Collapse unbranched paths in the graph, since the interior vertices are
redundant.  This should be relatively easy.  For example, branchless paths could
be extended from all vertices of degree 2.
\item Use a calculation called the A-statistic to compute the probabilitity that
each unbranched path contains unique sequence (i.e. should be traversed only 1
time).  The A-statistic for an unbranched path of length $\Delta$ that contains
$k$ internal vertices is $\Delta(n/G) - k \ln 2$, where $n$ is the total number
of reads and $G$ is the estimated genome size.  This step sounds
straightforward, but there are a couple things to consider.  First, when
collapsing unbranched paths in the previous step, the number of vertices in the
uncollapsed path must be remembered.  Secondly, the genome size must be
estimated, as it is not known.  And finally, reads that are fully contained in a
longer read should be mapped back into the graph to increase the accurancy of
the A-statistic calculations.
\item Calculate a traversal count for each edge in the graph to make the graph
Eulerian.  In the original paper, this is formulated as a minimum cost network
flow problem, so it should theoretically be possible to just set up the problem
and plug it into existing code, such as that in the LEMON library.  However,
since the actual problem is not actually a fully general minimum cost network
flow problem, there could eventually be some optimizations made by writing
stand-alone code.
\item Compute an Eulerian tour of the graph.  There may be several, but one
should be the original sequence.  I'm planning to do this using the stack-based
algorithm.
\item When doing evalutions of this algorithm, the last step is to compare the
assembly with the original genome.  This can be done with existing tools such as
the {\tt dnadiff} script included with the {\em MUMmer} sequence alignment
software.  This comparison will indicate how similar the assembled genome is to
the original genome.  Ideally, they should be identical.
\end{enumerate}

The actual implementation of this algorithm will be done in C++.  I will
initially implement the entire algorithm in one executable program, but I may
eventually split it into several executable problems if it makes sense to do so
(this is only possible once all the intermediate data structures can be
serialized and deserialized to external files).  Regardless of this, each step
of the algorithm will be split into one or more subroutines.  Logging will be
enabled by default.

\end{document}